A generalized definition of multidimensional item response theory parameters

### 2024-09-04 output/paper_draft/paper_M2PLM_parameters.Rmd

The representation of a latent space where the traits are correlated is more
accurate on a geometrical coordinate system with non-orthogonal axes
[@ackerman_multidimensional_2005-1] and (possibly) non-standard units,
often referred to as *general Cartesian coordinates* [@harman_modern_1970].
Such a space has the advantage that the basis vectors and the axes they yield
have a meaningful interpretation (albeit counterintuitive, as we will see)
in a multivariate statistical sense.
The general case of MIRT model estimation implies estimating possibly
non-standard latent dimensions and non-null covariances among them
(a more realistic case than an identity matrix).
Assuming orthonormality among traits is usually unrealistic, be it in the
cognitive [e.g., @reckase_multidimensional_2009] or the non-cognitive domain
[e.g., @thielmann_comparability_2022].
Moreover, certain applications impose the estimation of
non-identity latent covariance matrices (e.g., multigroup equating) even when
interpreting the latent dimensions as substantive constructs is not a goal
[for multidimensional linking and equating procedures,
see @reckase_linking_2009].
As we will argue, the usual definitions of the multidimensional difficulty and
discrimination parameters do not generalize to the non-orthonormal geometry
necessary to accurately represent correlated latent spaces.

>>>>

The representation of a latent space where the traits are correlated is more
accurate on a geometrical coordinate system with non-orthogonal axes
[@ackerman_multidimensional_2005-1] and (possibly) non-standard units,
often referred to as *general Cartesian coordinates* [@harman_modern_1970].
Such a space has the advantage that the basis vectors and the axes they yield
have a meaningful interpretation (albeit counterintuitive, as we will see)
in a multivariate statistical sense.
The general case of MIRT model estimation implies estimating possibly
non-standard latent dimensions and non-null covariances among them
(a more realistic case than an identity matrix).
Assuming orthonormality among traits is usually unrealistic, be it in the
cognitive [e.g., @reckase_multidimensional_2009] or the non-cognitive domain
[e.g., @thielmann_comparability_2022].
Moreover, certain applications impose the estimation of
non-identity latent covariance matrices (e.g., multigroup equating) even when
interpreting the latent dimensions as substantive constructs is not a goal
[for multidimensional linking and equating procedures,
see Chapter 9 in @reckase_multidimensional_2009].
As we will argue, the usual definitions of the multidimensional difficulty and
discrimination parameters do not generalize to the non-orthonormal geometry
necessary to accurately represent correlated latent spaces.
